An Embedding Framework for Consistent Polyhedral Surrogates

### 2022-01-07 poly-embed-journal/poly-embed-journal.tex
Redo proof of Claim 12 and formalize proof of Lemma 3

\begin{lemma}[\jessie{Lemma X}]\label{lem:X}
  Let $L: \R \to \reals^\Y_+$ be a minimizable loss with a polyhedral Bayes risk $\risk L$.
  Then $L$ has a finite representative set.
  Furthermore, letting $\Gamma = \prop{L}$, there exist finite sets
  $\V \subseteq \reals^\Y_+$ and
  $\Theta = \{\theta_v \subseteq \simplex \mid v\in\V\}$,
  both uniquely determined by $\risk{L}$ alone,
  such that
  \begin{enumerate}
  % \item There exists $\V = \V(\risk{L}) \subseteq \reals^\Y_+$, uniquely determined by $\risk{L}$ alone, such that $\V = L(\R^*)$.\label{item:X-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\V \subseteq L(\R')$.\label{item:X-rep-V}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $L(\R') = \V$.\label{item:X-min-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\Theta \subseteq \{\Gamma_r \mid r \in \R'\}$.\label{item:X-rep-Theta}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $\{\Gamma_r \mid r \in \R'\} = \Theta$.\label{item:X-min-Theta}
  \item Every representative set for $L$ contains a minimum representative set for $L$.\label{item:X-rep-contain-min}
  \item The set of full-dimensional level sets of $\Gamma$ is exactly $\Theta$.\label{item:X-full-dim}
  \item For any $r \in \R$, there exists $\theta \in \Theta$ such that $\Gamma_r \subseteq \theta$.\label{item:X-redundant}
  \item $L$ tightly embeds $\ell:\R'\to\reals^\Y_+$ if and only if $\ell$ is injective and $\ell(\R') = \V$.\label{item:X-tight-embed}
  \end{enumerate}
\end{lemma}

@@

\begin{claim}\label{claim:vface-subset-some-vstar-face}
	For any $v \in \V$, the face $F_{v}$ must be a subset of one of the $\V^*$ facets $F_{v^*}$ (as opposed to the vertical ones)
\end{claim}
\begin{proof}
	\jessiet{Can use a careful read}
	$\{F_{v^*}\}_{v^* \in \V^*}$ is exactly the set of facets of $G$ on $\reals_+^d$ by Theorem~\ref{thm:polyhedron-uniquely-gen-facets}.
	Any $k$-face is contained in a facet by Claim~\ref{claim:faces-contained-in-facets}.
	Since $\H = \cap (\H_{\V^*} \cup \H_\Y)$, we are done if $F_v$ is contained in a face of $\H_{\V^*}$ and need to consider the case where $F_v \in \H_\Y$.

	If $F_v$ is a subset of a vertical facet $F_y$, we would have $F_v = H_v \cap G \subseteq H_y \cap G$ for some $y \in \Y$.
	Expanding, this is $\{(x,g(x)) \mid \inprod{v}{x} = g(x)\} \subseteq \{(x,g(x)) \mid x_y = 0 \}$.
	While this can happen, it must be on the boundary of $\reals_+^d$, and must be at the intersection of $H_y$ and some other $H_{v^*}$ for $v^* \in \V^*$, since $g$ is continuous on $\reals^\Y_{+}$, and determined by $\V^*$ on $\reals^\Y_{++}$.
	\jessiet{Justification: $g$ is continuous on $\reals_+^\Y$, and there has to be a unique $v^*$ on $\reals^\Y_{++}$. More hand-wavy than I'd like}
\end{proof}

@@

Let $f_\V:\reals^\Y\to\reals_+\cup\{-\infty\}$ be a polyhedral concave function with $\dom(f_\V) = \simplex$.% such that $f_\V(x) = g_\V(x)$ for all $x \in \simplex$.
\begin{claim}
	We may write $f_\V(p) = \min_{v\in\V} \inprod{p}{v} + \delta(p \mid \simplex)$ for some finite set $\V \subset \reals^\Y_+$. %\raf{show you can do this using the $h$ and $\delta$ representation from Rockafellar -- argue that because $p\in\simplex$ we can write $\inprod{p}{b_i}-\beta_i = \inprod{p}{b_i-\beta_i\ones}$.}
\end{claim}
\begin{proof}
  We will think of $f_\V$ as defined $f_\V:\reals^\Y\to\reals_+\cup\{-\infty\}$  with $\dom(f_\V) = \simplex$.
  For $p \in \simplex$, we know $\sum_i p_i = 1$, and can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
  If $p \not \in \simplex$, then $f_\V(p) = -\infty$ and inner products are not used to compute $F$.
  Moreover, since $f_\V$ is polyhedral, it is finitely generated \citep[Proposition 19.1.2]{rockafellar1997convex} and can be written 
	\begin{align*}
	f_\V(p) &= h(p) - \delta(p \mid C)\\
		 &= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex)\\
		 &= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
	\end{align*}
\end{proof}

@@

\begin{proof}[Proofof Lemma~\ref{lem:X}]
Now this brings us to Lemma~\ref{lem:X}.
The framework in this appendix cues up this proof: any loss $L$ satisfying the assumptions of Lemma~\ref{lem:X} has some $g_{L(\R)} = \risk{L}_+$ as in this section that we can work with. 
For such a $g_{L(\R)}$, Lemma~\ref{lem:lemX1-rep-iff-subset-vectors} is exactly statement \eqref{item:X-rep-V}.
This immediately implies statement~\eqref{item:X-min-V}.
Moreover, Lemma~\ref{lem:lemX-3-rep-iff-FDLS-subsets} is exactly statement~\eqref{item:X-rep-Theta}, and again statement~\eqref{item:X-min-Theta} immediately follows.
Statement~\eqref{item:X-rep-contain-min} is a corollary of the existence of a finite representative set, as shown in Corollary~\ref{cor:finite-rep-set}.
Again, Statement~\eqref{item:X-full-dim} is exactly Lemma~\ref{lem:fdls-exactly-theta}.
Statement~\eqref{item:X-redundant} is exactly Lemma~\ref{lem:any-levelset-contained-in-minlevelset}.
Finally, Statement~\eqref{item:X-tight-embed} follows as a corollary of statement~\eqref{item:X-min-V} and Corollary~\ref{cor:tight-embed-min-rep}.
\end{proof}

>>>>

\begin{restatable}{lemma}{lemmaX}[\jessie{Lemma X}]\label{lem:X}
  Let $L: \R \to \reals^\Y_+$ be a minimizable loss with a polyhedral Bayes risk $\risk L$.
  Then $L$ has a finite representative set.
  Furthermore, letting $\Gamma = \prop{L}$, there exist finite sets
  $\V \subseteq \reals^\Y_+$ and
  $\Theta = \{\theta_v \subseteq \simplex \mid v\in\V\}$,
  both uniquely determined by $\risk{L}$ alone,
  such that
  \begin{enumerate}
  % \item There exists $\V = \V(\risk{L}) \subseteq \reals^\Y_+$, uniquely determined by $\risk{L}$ alone, such that $\V = L(\R^*)$.\label{item:X-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\V \subseteq L(\R')$.\label{item:X-rep-V}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $L(\R') = \V$.\label{item:X-min-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\Theta \subseteq \{\Gamma_r \mid r \in \R'\}$.\label{item:X-rep-Theta}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $\{\Gamma_r \mid r \in \R'\} = \Theta$.\label{item:X-min-Theta}
  \item Every representative set for $L$ contains a minimum representative set for $L$.\label{item:X-rep-contain-min}
  \item The set of full-dimensional level sets of $\Gamma$ is exactly $\Theta$.\label{item:X-full-dim}
  \item For any $r \in \R$, there exists $\theta \in \Theta$ such that $\Gamma_r \subseteq \theta$.\label{item:X-redundant}
  \item $L$ tightly embeds $\ell:\R'\to\reals^\Y_+$ if and only if $\ell$ is injective and $\ell(\R') = \V$.\label{item:X-tight-embed}
  \end{enumerate}
\end{restatable}

@@

\begin{claim}\label{claim:vface-subset-some-vstar-face}
	For any $v \in \V$, the face $F_{v} \subseteq F_{v^*}$ for some $v^* \in \V^*$. %must be a subset of one of the $\V^*$ facets $F_{v^*}$ (as opposed to the vertical ones)
\end{claim}
\begin{proof}
	As $\hyp(g_{\V^*})$ is polyhedral, each of its faces are convex polytopes, and is also a face of some facet of $\hyp(g_{\V^*})$; these facets are defined by $\V^*$ and $\Y$ (Corollary 6).

	It then suffices to show that if $F_v$ is a face of the facet $F_y$ for some $y\in \Y$, then it must also be a face of $F_{v^*}$ for a $v \in \V^*$; equivalently, $F_v$ is not a facet, and thus $F_v \neq F_y$.
	Recall from the definition of a face that $F_v = G \cap H_{(v,-1)}$ and $F_y = G \cap H_{(e_y, 0)}$.
	As facets are (uniquely) determined by halfspaces, and $F_y$ is a facet, we either have $F_v \neq F_y$ (in particular, if $F_v$ is a facet) or $F_v \subsetneq F_y$ (if $F_v$ is not a facet).
	In both cases, we have $F_v \neq F_y$, and the result follows.
	\jessie{I'm not sure if this is clear, but this is a new approach to the proof that I generally like better. I can clarify if the wording is not clear though.}

	\jessiet{Old version in comments}
	%Commented out 7 Jan 2022
%	\jessiet{Can use a careful read}
%	$\{F_{v^*}\}_{v^* \in \V^*}$ is exactly the set of facets of $\hyp(g_{\V^*})$ on $\reals_+^d$ by Theorem~\ref{thm:polyhedron-uniquely-gen-facets}.
%	Any $k$-face is contained in a facet by Claim~\ref{claim:faces-contained-in-facets}.
%	Since $\H = \cap (\H_{\V^*} \cup \H_\Y)$, we are done if $F_v$ is contained in a face of $\H_{\V^*}$ and need to consider the case where $F_v \in \H_\Y$.
%	
%	If $F_v$ is a subset of a vertical facet $F_y$, we would have $F_v = H_v \cap G \subseteq H_y \cap G$ for some $y \in \Y$.
%	Expanding, this is $\{(x,g(x)) \mid \inprod{v}{x} = g(x)\} \subseteq \{(x,g(x)) \mid x_y = 0 \}$.
%	While this can happen, it must be on the boundary of $\reals_+^d$, and must be at the intersection of $H_y$ and some other $H_{v^*}$ for $v^* \in \V^*$, since $g$ is continuous on $\reals^\Y_{+}$, and determined by $\V^*$ on $\reals^\Y_{++}$.
%	\jessiet{Justification: $g$ is continuous on $\reals_+^\Y$, and there has to be a unique $v^*$ on $\reals^\Y_{++}$. More hand-wavy than I'd like}
\end{proof}

@@

Let $f_\V:\reals^\Y\to\reals_+\cup\{-\infty\}$ be a polyhedral concave function with $\dom(f_\V) = \simplex$.% such that $f_\V(x) = g_\V(x)$ for all $x \in \simplex$.
\begin{claim}\label{claim:f-min-affine}
	We may write $f_\V(p) = \min_{v\in\V} \inprod{p}{v} + \delta(p \mid \simplex)$ for some finite set $\V \subset \reals^\Y_+$. %\raf{show you can do this using the $h$ and $\delta$ representation from Rockafellar -- argue that because $p\in\simplex$ we can write $\inprod{p}{b_i}-\beta_i = \inprod{p}{b_i-\beta_i\ones}$.}
\end{claim}
\begin{proof}
  We will think of $f_\V$ as defined $f_\V:\reals^\Y\to\reals_+\cup\{-\infty\}$  with $\dom(f_\V) = \simplex$.
  For $p \in \simplex$, we know $\sum_i p_i = 1$, and can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
  If $p \not \in \simplex$, then $f_\V(p) = -\infty$ and inner products are not used to compute $F$.
  Moreover, since $f_\V$ is polyhedral, it is finitely generated \citep[Proposition 19.1.2]{rockafellar1997convex} and can be written 
	\begin{align*}
	f_\V(p) &= h(p) - \delta(p \mid C)\\
		 &= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex)\\
		 &= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
	\end{align*}
\end{proof}

@@

Now this brings us to Lemma~\ref{lem:X}.
The framework in this appendix cues up this proof: any loss $L$ satisfying the assumptions of Lemma~\ref{lem:X} has some $g_{L(\R)} = \risk{L}_+$ as in this section that we can work with.

\lemmaX*
\begin{proof}
%As in this appendix, we consider $\V = L(\R)$ such that $g_\V = \risk L_+$.
Consider $\risk L = f_\V$ for a finite set $\V$ by Claim~\ref{claim:f-min-affine}.
There is a polyhedral concave function $g_\V$ on $\reals^\Y_+$ matching $f_\V$ on $\simplex$ by Corolary~\ref{cor:f-matches-g-on-simplex}.
Moreover, consider $g_\V = \risk L_+$, and observe that $\risk L_+$ matches $\risk L$ on $\simplex$ as well.
By Corollary~\ref{cor:unique-set-loss-vectors-defining-facets}, we then have a finite set $\V^*$ of smallest cardinality such that $f_\V = f_{\V^*}$ and $g_\V = g_{\V^*}$.
Consider $\R^* \subseteq \R$ such that $\V^* = L(\R^*)$.
First, observe that $\R^*$ is representative for $L$ as a corollary of Claim~\ref{claim:projected-faces-cover-RY-iff-representative}.
Moreover, consider the follow set of level sets of $f_{\V^*}$, $\Theta = \{\theta_v \mid v \in \V^*\}$.

Now that we have the preliminaries, consider the itemized statements.
For $f_{\V^*}$, Lemma~\ref{lem:lemX1-rep-iff-subset-vectors} is exactly statement \eqref{item:X-rep-V}.
This immediately implies statement~\eqref{item:X-min-V}.
Moreover, Lemma~\ref{lem:lemX-3-rep-iff-FDLS-subsets} is exactly statement~\eqref{item:X-rep-Theta}, and again statement~\eqref{item:X-min-Theta} immediately follows.
Statement~\eqref{item:X-rep-contain-min} is a corollary of the existence of a finite representative set, as shown in Corollary~\ref{cor:finite-rep-set}.
Again, Statement~\eqref{item:X-full-dim} is exactly Lemma~\ref{lem:fdls-exactly-theta}.
Statement~\eqref{item:X-redundant} is exactly Lemma~\ref{lem:any-levelset-contained-in-minlevelset}.
Finally, Statement~\eqref{item:X-tight-embed} follows as a corollary of statement~\eqref{item:X-min-V} and Corollary~\ref{cor:tight-embed-min-rep}.
\end{proof}
